Correct and Secure Web Programming using Dependent Types and Embedded Domain-Specific Languages

### 2013-07-26 paper/main.tex
More work on paper

\titlebanner{Draft for submission to IFL 2013}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.
\title{Correct and Secure Web Programming using Dependent Types and
Embedded Domain-Specific Languages}
%\subtitle{Using Embedded Domain-Specific Languages to Improve Confidence in Web Applications}

@@

\begin{abstract}

\end{abstract}

@@

@@

With this proof, we may then use the \texttt{with} construct, inspired by \textit{views} in Agda \cite{viewfromleft}, to rewrite the arguments on the left hand side. This allows us to construct a function which, given the handler stored in the list of handlers, the data required to construct the function type and the \texttt{MkHandlerFnTy} deserialised from the form, determines whether the two \texttt{MkHandlerFnTy}s are decidably equal. If so, we may rewrite this on the left hand side, and therefore demonstrate that the recorded function may also be used to handle the form data. If not, the computation is abandoned and an error is shown. % I gave it a shot, right? :P
{\small
\begin{verbatim}
checkFunctions : (reg_fn_ty : MkHandlerFnTy) -> 
                 (frm_fn_ty : MkHandlerFnTy) -> 
                 mkHandlerFn reg_fn_ty -> 
                 Maybe (mkHandlerFn frm_fn_ty)
checkFunctions reg_ty frm_ty reg_fn with 
                             (decEq reg_ty frm_ty)
  checkFunctions frm_ty frm_ty reg_fn | Yes refl = Just reg_fn
  checkFunctions reg_ty frm_ty reg_fn | No _ = Nothing
\end{verbatim}}
We may then parse the arguments according to the types specified by the handler function, and then apply the arguments to the handler function through the use of partial application. Finally, we may run the handler function, ensuring that all updates made to the CGI state are propagated.
\section{Extended Example: Message Board}
\section{Related Work}
\section{Conclusions}
\subsection{Further Work}
% Cryptography bindings would be really good.
% Integration with a web server instead of doing everything over CGI
% Improvements to form handling system

@@

\acks
% SICSA / EPSRC (grant number?)
% #idris
% Idris contributors, in particular Melissa for the SQLite bindings and whoever wrote Network.Cgi

>>>>

\titlebanner{Draft for submission to IFL 2013}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

\title{Correct and Secure Web Programming using Dependent Types and Embedded Domain-Specific Languages}
%\subtitle{Using Embedded Domain-Specific Languages to Improve Confidence in Web Applications}

@@

\begin{abstract}
Dependently-typed languages allow types to be used to impose arbitrarily-specific constraints on programs, in turn allowing for easier reasoning about the operation of such programs. Stronger specifications do however bring with them the disadvantage that it becomes increasingly difficult to write programs that are accepted by the type checker, and additional proofs may have to be specified by a user.

Research into embedded domain-specific languages attempt to address this problem by introducing a layer of abstraction over the underlying types, allowing for domain-specific code to be written in high-level languages which use dependent types to enforce certain invariants without requiring complex proofs to be specified.

In this paper, we apply this technique to web programming, showing that embedded domain-specific languages can be constructed to enforce resource usage protocols associated with common web operations such as CGI, database access and session handling. Additionally, we show that this technique may be employed to allow the creation and handling of web forms which retain type information, reducing the scope for programmer error and attacks which exploit the loss of type information, such as SQL injection.
\end{abstract}

@@

In order to decrease unnecessary boilerplate code in user applications, we provide functions which abstract out unnecessary parts of this process. In order to do this, we define the algebraic data type \texttt{DBVal}, which is a tagged union over simple primitive types:
{\small
\begin{verbatim}
data DBVal = DBInt Int
           | DBText String
           | DBFloat Float
           | DBNull
\end{verbatim}
}
We also make use of the \texttt{ResultSet} type, which is a list of returned database rows.
\begin{verbatim}
ResultSet : Type
ResultSet = List (List DBVal)
\end{verbatim}
Using these, we may implement a function, \texttt{ExecuteSelect}, which, given a query, a list of variables to bind and their associated indices within the query, and a function which is used to extract information out of each returned database row, returns either a \texttt{ResultSet} or an error.
{\small
\begin{verbatim}
executeSelect : String ->
                String -> 
                List (Int, DBVal) -> 
                (Eff m [SQLITE 
                  (SQLiteRes PreparedStatementExecuting)
                   ] 
                 (List DBVal)
                ) -> 
                Eff m [SQLITE ()] (Either String ResultSet)
\end{verbatim}
}
% =================================================

@@

With this proof, we may then use the \texttt{with} construct, inspired by \textit{views} in Agda \cite{viewfromleft}, to rewrite the arguments on the left hand side. This allows us to construct a function which, given the handler stored in the list of handlers, the data required to construct the function type and the \texttt{MkHandlerFnTy} deserialised from the form, determines whether the two \texttt{MkHandlerFnTy}s are decidably equal. If so, we may rewrite this on the left hand side, and therefore demonstrate that the recorded function may also be used to handle the form data. If not, the computation is abandoned and an error is shown. % I gave it a shot, right? :P
{\small
\begin{verbatim}
checkFunctions : (reg_fn_ty : MkHandlerFnTy) -> 
                 (frm_fn_ty : MkHandlerFnTy) -> 
                 mkHandlerFn reg_fn_ty -> 
                 Maybe (mkHandlerFn frm_fn_ty)
checkFunctions reg_ty frm_ty reg_fn with 
                             (decEq reg_ty frm_ty)
  checkFunctions frm_ty frm_ty reg_fn | Yes refl = Just reg_fn
  checkFunctions reg_ty frm_ty reg_fn | No _ = Nothing
\end{verbatim}}
We may then parse the arguments according to the types specified by the handler function, and then apply the arguments to the handler function through the use of partial application. Finally, we may run the handler function, ensuring that all updates made to the CGI state are propagated.
\section{Extended Example: Message Board}
\section{Related Work}
\section{Conclusions}
Dependently-typed languages have, for a long time, remained relatively unused outside of academia. Advances in embedded domain-specific languages have allowed many of the advantages of such languages, including the additional guarantees given by the successful compilation of a program (assuming the use of specific enough tyeps) to be adopted without the additional proofs traditionally required by users of such languages.

We believe we have demonstrated that embedded domain-specific languages can be used to make use of the power of dependently-typed languages without incurring the traditional overheads of 
\subsection{Further Work}
We have shown that embedded domain-specific languages using dependent types and algebraic effects can be used to increase confidence in web applications by providing additional static guarantees about runtime behaviour, but much more can be done using the same approach.

There are firstly many other applications which make use of specific resource usage protocols, for example popular libraries such as \textit{libgcrypt} \cite{libgcrypt}. Applying a similar approach would allow for sensitive programs requiring cryptographic routines to be written using a language with full dependent types, in turn adding an extra layer of confidence to their security.

Whilst the use of CGI allows for experimenting with the use of dependent types in a real-world scenario such as web programming, there remain practical considerations about its scalability, as a separate process must be created for each individual request request. Ultimately, we would like to create a web server written in \idris{}, which would make more efficient usage of resources.

Dependently-typed languages provide great promise for the construction of secure and correct programs. Through the use of embedded domain-specific languages, we hope that
% Cryptography bindings would be really good.
% Integration with a web server instead of doing everything over CGI
% Improvements to form handling system

@@

\acks
% SICSA / EPSRC (grant number?
% #idris
% Idris contributors, in particular Melissa for the SQLite bindings and whoever wrote Network.Cgi
